D-Hammer: Efficient Equational Reasoning for Labelled Dirac Notation

### 2025-04-10 05tool_eval.tex
.

\subsubsection*{Quantum circuits}
Quantum circuits is a prominent model of quantum computation. This
model is adopted by numerous tools, which can evaluate, optimize, or
prove equivalence of quantum circuits. These tools are based on a
variety of approaches, based on ZX-calculus~\cite{9868772}, or
decision diagrams\cite{10.1145/3394885.3431590}, or other formalisms
discussed in related work. In general, these tools are aggressively
optimized to achieve scalability.

Quantum circuits can also be described as unitary operators in Dirac
notation. Therefore, D-Hammer can check the equivalence of quantum
circuits through their Dirac notation representations. Although it is
not the intended application of D-Hammer, we include an evaluation of
D-Hammer on some simple examples. An evaluation of some examples is
given in~\Cref{fig: circuit eval}, and~\Cref{fig: QC-5} shows one of
them. As expected, D-Hammer does not perform well in comparison with
specialized tools: it is always clearly outperformed on small quantum
circuit examples, and it always times out on larger quantum circuit
examples. It would be of interest to make D-Hammer more competitive on
quantum circuits in future work.

>>>>

\subsubsection*{Quantum circuits}
Quantum circuits is a prominent model of quantum computation. This
model is adopted by numerous tools, which can evaluate, optimize, or
prove equivalence of quantum circuits. These tools are based on a
variety of approaches, based on ZX-calculus~\cite{9868772}, or
decision diagrams~\cite{10.1145/3394885.3431590}, or other formalisms
discussed in \Cref{sec:relwork}. In general, these tools are aggressively
optimized to achieve scalability.

Quantum circuits can also be described as unitary operators in Dirac
notation. Therefore, D-Hammer can check the equivalence of quantum
circuits through their Dirac notation representations. Although it is
not the intended application of D-Hammer, we include an evaluation of
D-Hammer on some simple examples. An evaluation of some examples is
given in~\Cref{fig: circuit eval}, and~\Cref{fig: QC-5} shows one of
them. As expected, D-Hammer does not perform well in comparison with
specialized tools: it is always outperformed, often by several order
of magnitude, on small quantum circuit examples, and it always times
out on larger quantum circuit examples. In the future it would be of
interest to make D-Hammer more competitive on quantum circuits by
adopting some of the aggressive optimizations for other works.
